Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-14 paper/main.tex
position labels to avoid interference with plots
include hint for discussion around loss functions of semantic segmentation
rename; relabel
add tableofcontents to create overivew for iterations
write section on fundamental assumptions

@@

%###################################################################################################
\section{Modeling Prerequesites}
\begin{enumerate}
    \item represent bacteria as polygon with vertices $\vec{v}_i$
    \item Assumes that these vertices are not fundamental but rather a discretization of the actual
        problem
\end{enumerate}

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-004000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-006000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-008000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        Progression of generated masks with estimated and exact positions.
        Comparison of masks across generations. Compare estimated with exact rod lengths.
    }
    \label{fig:benchmarking-extraction-algorithm}
\end{figure}

@@

% --------------------------------------------------------------------------------------------------
\subsection{Constructing a Cost Function}

@@

@@

%###################################################################################################
\section{Parameter Estimation}
\label{sec:supplement-parameter-profiles}
%TODO {explain profiles}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-5.pdf}
    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-5.pdf}%
    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement}
\end{figure}

>>>>

\tableofcontents

@@

%###################################################################################################
\section{Fundamental Assumptions on the Spatial Representation of Rod-Shaped Bacteria}
% TODO CITATIONS
This work makes use of a few but powerfull assumptions concerning the mathematical representation of
the cells in question.
In order to describe the spatial dynamics of individual agents, we need to define their position,
shape and how they interact with each other.
In order to provide as much flexibility as possible when constructing new models, we restrict
ourselves to assumptions concerning the position such that researchers can freely experiment with
various shapes and interaction types.
Furthermore, it is very important to strike a good balance between the restrictiveness of our
assumptions and generality in order to provide meaningful results and still be able to apply our
approach to a broader class of models.
Taking into account the natural shape of rod-like bacteria, it is clear that they can not be
represented with a singular point $\vec{v}$ in space.
We could extend this positional definition with a vector $\vec{w}$, indicating the direction of
extension but this would restrict ourselves to fully straight rod-shaped bacteria.
However, multiple experiments have shown that MReB is responsible for the growth and maintenance of
the elongated shape of bacteria~\cite{Erickson2001}.
By means of this protein, rod-shaped bacteria can bend elastically on short time-scales and
deform plastically over longer periods of time.
Our goal is to provide a model which can encapsulate these effects.
Thus, we extend our model from a single point to multiple vertices which are part of a 1-dimensional
polygon with two endpoints $\vec{v}_i$.
In essence, we discretize the (possibly non-straight) rod of the agent.
Figure~\ref{fig:mechanics-bacterium} shows a visualization of this approach.
The angles $\alpha_i$ between vertices $\vec{v}_{i-1},\vec{v}_i$ can be used when modeling the
dynamics of these bacteria.

With these fundamental thoughts layed out, we retain the flexibility to represent many models such
as varying shapes since we do not have to make any assumptions on the type of interactions between
two bacteria.
Even concerning the shape of bacteria, this assumption still allows us to mdoel elliptical cells by
modifying the interaction type as well as other effects such as polarity of
interactions~\cite{Duvernoy2018}.
Furthermore, we also did not have to assume any particular mechanical dynamics of the agents, only
their spatial representation, thus leaving in particular room for plastical or elastic deformations
of the rod and other more intricate mechanisms.

\begin{enumerate}
    \item represent bacteria as polygon with vertices $\vec{v}_i$
    \item Assumes that these vertices are not fundamental but rather a discretization of the actual
        problem
\end{enumerate}

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-004000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-006000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-008000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        Progression of generated masks with estimated and exact positions.
        Comparison of masks across generations. Compare estimated with exact rod lengths.
    }
    \label{fig:benchmarking-extraction-algorithm}
\end{figure}

@@

% --------------------------------------------------------------------------------------------------
\subsection{Constructing a Loss Function}

@@

\paragraph{Comparison with Semantic Segmentation}
\label{paragraph:comparison-semantic-segmentation}
\cite{Jadon2020}

@@

%###################################################################################################
\section{Parameter Estimation}
\label{sec:supplement-parameter-profiles}
%TODO {explain profiles}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-5.pdf}
    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-growth-rates}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-5.pdf}%
    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-radii}
\end{figure}
